Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-08 arxiv/proof_examples.tex
revise

Now we are ready to prove equation \eqref{hatw_add1}. First, using equations \eqref{op_uplow} and  \eqref{h'(A)1}, we obtain that for $1+d^2/\kappa^2 \le x\le 1+\delta$,
$$h'(x)\ge \frac{2(x-1)\kappa^2 - d^2}{x^3} \cdot p^{-1} \tr\left[\left([(X^{(2)})^\top X^{(2)}]^{-1} + x^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right] \gtrsim p\left(d^2 + |x-1|\kappa^2\right).$$
Thus we get that if $1+2d^2/\kappa^2 +p^{-1/4+c}\le x\le 1+\delta$,
\begin{align*}
h(x)-h(1)\ge h(x)-h(1+d^2/\kappa^2) \ge \int_{1+d^2/\kappa^2}^x h'(x)\dd x \gtrsim p\kappa^2 \cdot |x-1|^2  \gtrsim p\frac{d^4}{\kappa^2} + p^{-1/2+2c}\cdot p\kappa^2 .
\end{align*}
On the other hand, by Claim \ref{claim_largedev1} we have that
$$ g(x)=h(x)+\sigma^2 (n_1+n_2-p)+\OO \left(p^{1/2+c}\left(\sigma^2 +\kappa^2+d^2 \right)\right)$$
with high probability. Combining the above two estimates, we obtain that under the conditions $\sigma^2 \lesssim  \kappa^2$ and $d^2 \le p^{-c_{\varphi}}{\kappa^2}$,
$$g(x)-g(1) = h(x)-h(1) + \OO \left(p^{1/2+c}\left(\sigma^2 +\kappa^2+d^2 \right)\right) >0$$
with high probability, i.e. $x$ cannot be the minimizer of $g$. 
%Then if we choose the constant $c$ in Claim \ref{claim_largedev1} such that $e_1<2e_2$, we have 
%$$ h(x)-h(1)\gtrsim  p^{-1/2+2e_2}\cdot p\kappa^2 \gg p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right),$$ where we use the assumption that $\sigma^2$ and $d^2$ are both at most $\OO(\kappa^2)$, which implies the inequality \eqref{gA>g1}. 
Second, using equations \eqref{op_uplow} and  \eqref{h'(A)2}, we obtain that for $1-\delta \le x \le 1-d^2/\kappa^2$,
$$- h'(x)\ge  [2(1-x)\kappa^2 - xd^2] \cdot p^{-1} \tr\left[\left(x^2[(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right]  \gtrsim p\left(d^2 + |x-1|\kappa^2\right).$$
Then with a similar argument as above, we can get that if $1-2d^2/\kappa^2 - p^{-1/4+c}\le x\le 1-\delta$, then 
$$ h(x)-h(1) \gtrsim p\frac{d^4}{\kappa^2} +p^{1/2+2c} \kappa^2, $$
%under equation \eqref{choiceofpara0} \HZ{I don't see which condition we want to use here?}, 
which again implies that $g(x)>g(1)$, i.e. $x$ cannot be the minimizer of $g$. In sum, we obtain that the minimizer $\hat x$ must satisfy equation \eqref{hatw_add1}.
%Next we prove the following estimate on the optimizer $\hat A$: with high probability,
%%\begin{lemma}
%%For the isotropic model, we have
%\be\label{hatv_add1}
%|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2 + \e_0 /2+ 2\e}.
%\ee
%%\end{lemma}
%%\begin{proof}
% In fact, from the proof of Claim \ref{lem_hat_v} above, one can check that if $C\cal E \le |w-1| \le 2C\cal E$ for a large enough constant $C>1$, then $|h'(w)|\gtrsim \sigma^2$. Moreover, under equation \eqref{choiceofpara} we have
%$$h(w) =\OO(\sigma^2),\quad \text{for}\quad   |w-1|\le 2C\cal E.$$
%Thus we obtain that for $|w-1|\ge 2C\cal E$,
%$$\left|h(w) - h(\hat w)\right|\ge |h(w)-\min\{h(w_1),h(w_2)\}|\gtrsim \sigma^2 \cal E \gtrsim \cal E \cdot h(\hat w),$$
%which leads to $g(w) > g(\hat w)$ with high probability by equation \eqref{boundv-w}. Thus $w$ cannot be a minimizer of $g(v)$, and we must have $|\hat v - 1|\le 2C\cal E$. %Together with equation \eqref{hatw_add1}, we conclude equation \eqref{hatv_add1}.
%%\end{proof}
%
%{\cor require $\kappa^2\gg d^2 + p^{-1/2+c}\sigma^2$}
\end{proof}

\paragraph{Part $3$: Applying Theorem \ref{thm_main_RMT}.}
We provide the predication loss using the optimal $\hat B$ in equation \eqref{eq_Bhat_2task} and optimal $\hat x=\hat A_1/\hat A_2$:
\begin{align}
L(\hat{\beta}_2^{\MTL}) &=\left|\Sigma_2^{1/2} \left( \hat B \hat A_2 - \beta_2\right)\right| \nonumber\\
&=  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat x)^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat x\beta_1-\hat x^2\beta_2)+ (X^{(2)})^\top \epsilon^{(2)} + \hat x   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2.\label{Lbeta_HPS}
\end{align}
%where we denoted $$\hat \Sigma:= \hat A^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$$
To illustrate the idea of bias-variance decomposition, we assume that $\hat x$ is independent of $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$ for now. Then we can calculate the partial expectation of $L(\hat{\beta}_2^{\MTL})$ over the randomness in $\epsilon_1$ and $\epsilon_2$:
 \begin{align}
	&\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2,\beta_1,\beta_2} \nonumber \\
&	=  \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)} (\hat A \beta_1 - \hat A^2\beta_2)}^2 \nonumber\\ %\label{eq_bias_2task} 
			& + \sigma^2\cdot \bigtr{ X^{(2)} \hat\Sigma^{-1}\Sigma^{(2)}\hat\Sigma^{-1} (X^{(2)})^\top}+ \sigma^2\cdot \hat A^2 \bigtr{ X^{(1)} \hat\Sigma^{-1}\Sigma^{(2)}\hat\Sigma^{-1} (X^{(1)})^\top}\nonumber\\
			%+\sigma^2\cdot \bigtr{\Sigma_2(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } \nonumber\\ 
			&= \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)} (\hat A \beta_1 - \hat A^2\beta_2)}^2 +\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma^{-1}\right).\nonumber
			%\label{eq_var_2task}
\end{align}
This gives a bias-variance decomposition of the MTL prediction loss: the first term on the right-hand side is the bias term, while the second term is the variance term. For the random-effects model, we further take the partial expectation over $\beta_1$ and $\beta_2$: 
 \begin{align}
	\exarg{\epsilon^{(1)}, \epsilon^{(2)},\beta_1,\beta_2}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2}&	=  \left[(\hat A-1)^2 \kappa^2 + (\hat A^2+1)\frac{d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma^{-1} \Sigma^{(2)} \hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)} }\nonumber \\
	&+\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma^{-1}\right).
			\label{eq_var_2task}
\end{align}
One can compare the above expression with the predication loss for STL, that is,
\begin{align}
	\exarg{\epsilon_2}{L(\hat{\beta}_2^{\STL}) \mid X_2} = \sigma^2 \cdot \bigtr{\Sigma_2 (X_2^{\top} X_2)^{-1}}, \label{eq_var_stl}
\end{align}
we observe that while the bias of $\hat{\beta}_2^{\MTL}$ is always larger than that of $\hat{\beta}_2^{\STL}$, which is zero, the variance of $\hat{\beta}_2^{\MTL}$ is always lower than that of $\hat{\beta}_2^{\STL}$.\footnote{To see why this is true, we apply the Woodbury matrix identity over the second term on the right-hand side of equation \eqref{eq_var_2task} and use the fact that for the product of two PSD matrices, its trace is always nonnegative.}
In other words, training both tasks together helps predict the target task by reducing variance while incurring a bias.
Therefore, whether multi-task learning outperforms single-task learning is determined by the bias-variance decomposition!

@@

In general, for $\hat A$ as  the minimizer of $g(x)$, it depends on $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$. However, using the concentration estimates in Lemma \ref{largedeviation}, we can still obtain the expression on the right-hand side of \eqref{eq_var_2task} up to a small error. In particular, our proof will not use Lemma \ref{lem_hat_v}.

%control the difference between $g(x)$ and its partial expectation. 
 \begin{claim}\label{claim_largedev2}
 In the setting of Proposition \ref{cor_MTL_loss}, if $\hat A=\OO(1)$, then we have that with high probability,
$$\left| L(\hat{\beta}_2^{\MTL})- \cal L(\hat A)\right| \le p^{-1/2+e} \left(\sigma^2 +\kappa^2 +d^2 \right),$$
for any small constant $e>0$. Here we introduced the function
\begin{align*}
	\cal L(\hat A)	:=  \left[(\hat A-1)^2 \kappa^2 + (\hat A^2+1)\frac{d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma^{-1} \Sigma^{(2)} \hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)} } +\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma^{-1}\right).
\end{align*}
 \end{claim}
 \begin{proof}
 The proof of this claim is very similar to the one for Claim \ref{claim_largedev1}, where the only difference is that $\hat A$ now may depend on $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$. In this case, we can still use  Lemma \ref{largedeviation} to conclude the proof because $\hat A$ can be pulled out as a coefficient. Here we only give an outline of the proof without writing down all the details. Recall that $\beta_0$ is the shared component of $\beta_1$ and $\beta_2$ with i.i.d. Gaussian entries of mean zero and variance $p^{-1}\kappa^2$. Moreover, we denote the task-specific components by $\wt\beta_1$ and $\wt\beta_2$, whose entries i.i.d. Gaussian random variables of mean zero and variance $p^{-1} d^2 / 2$. Then we write $L(\hat{\beta}_2^{\MTL}) $ in \eqref{Lbeta_HPS} as:
 \begin{align}
L(\hat{\beta}_2^{\MTL})  =&  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat A -\hat A^2)\beta_0+(X^{(1)})^\top X^{(1)} \hat A\wt \beta_1 - (X^{(1)})^\top X^{(1)}  \hat A^2\wt \beta_2 \right] \right. \\
&\left. + (\Sigma^{(2)})^{1/2}\hat \Sigma^{-1}\left[ (X^{(2)})^\top \epsilon^{(2)} + \hat A   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2. \label{expand_15}
\end{align}
As in \eqref{expand_6}, we can expand this expression into the sum of 15 terms, and bound each term as in \eqref{eq_hA1}-\eqref{eq_hA56}. For example, for the main term $\hat A^2 \wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(1)})^\top X^{(1)}\wt\beta_1$, using 
Lemma \ref{largedeviation} and \eqref{op_X12}-\eqref{op_Sig1}, we can obtain the following estimate with high probability for any small constant $e>0$:
\begin{align*}
&\hat A^2 \left|\wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(1)})^\top X^{(1)}\wt\beta_1 - d^2 \cdot p^{-1}\bigtr{(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(1)})^\top X^{(1)}}\right| \\
&\le p^e\cdot p^{-1}d^2 \cdot \left\| (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(1)})^\top X^{(1)} \right\|_F \\
&\le p^{-1/2+\e}d^2 \cdot  \left\| (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(1)})^\top X^{(1)} \right\|  \lesssim p^{-1/2+\e}d^2.
\end{align*} 
For the cross term $\hat A \wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(2)})^\top  \epsilon^{(2)}$, using Lemma \ref{largedeviation} and \eqref{op_X12}-\eqref{op_Sig1}, we can obtain the following estimate with high probability for any small constant $e>0$:
\begin{align*}
&|\hat A |\cdot \left|\wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(2)})^\top  \epsilon^{(2)}\right| \le p^e \cdot \sigma\sqrt{p^{-1}d^2} \cdot \|(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(2)})^\top  \|_F \\
&\lesssim p^{c}\sigma d \cdot  \|(X^{(1)})^\top X^{(1)} \hat \Sigma^{-1}  \Sigma^{(2)} \hat \Sigma^{-1}  (X^{(2)})^\top  \| \lesssim p^{-1/2+e}\sigma d\le p^{-1/2+e}(\sigma^2+ d^2).
\end{align*}
The rest of the terms in the expansion of \eqref{expand_15} can be bounded in the same way, and we omit the details.
 \end{proof}

In the specific setting in Proposition \ref{cor_MTL_loss}, we can further simplify $\cal L(\hat A)$ using Claim \ref{lem_hat_v} and $\Sigma^{(1)}=\Sigma^{(2)}=\id_p$.

@@

&\left|\cal L(\hat A)- d^2\cdot {p}^{-1}\tr\left[\Sigma(1)^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \bigtr{\Sigma(1)^{-1}  }\right| \nonumber\\
&\lesssim  \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2e}\kappa^2 +  p^{-1/4+e} (\sigma^2 +d^2) , \label{simple1}

@@

for any small constant $e>0$, where we recall that $\Sigma(1)= (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)}$.

@@

Using Lemma \ref{lem_hat_v} and Fact \ref{lem_minv} (iii), we can obtain that with high probability,

@@

&\|\Sigma(1)^{-1}-\hat\Sigma^{-1}\| \le |\hat A-1|\|\Sigma(1)^{-1}\| \| (X^{(1)})^\top X^{(1)}\|\|\hat\Sigma^{-1}\| \label{est111}\\
&\lesssim \left(\frac{d^2}{\kappa^2} + p^{-1/4+e}\right)\frac{(\sqrt{n_1}+\sqrt{p})^{2}+n_1p^{-c_\varphi}}{ \left[(\sqrt{n_2}-\sqrt{p})^{2}-n_2p^{-c_\varphi}  + (\sqrt{n_1}-\sqrt{p})^{2}-n_1p^{-c_\varphi}\right]^2  }\lesssim p^{-1}\left(\frac{d^2}{\kappa^2} + p^{-1/4+e}\right),\nonumber

@@

for any small constant $e>0$. With similar arguments, we can get that with high probability,

@@

&\left\|\left(\Sigma(1)^{-2}-\hat\Sigma^{-2}\right)\left((X^{(1)})^\top X^{(1)} \right)^2\right\| \lesssim  \frac{d^2}{\kappa^2} + p^{-1/4+e} .

@@

Moreover, using Fact \ref{lem_minv} (iii) we can bound that with high probability,

>>>>

Now we are ready to prove equation \eqref{hatw_add1}. First, using equations \eqref{op_uplow} and  \eqref{h'(A)1}, we obtain that for $1+d^2/\kappa^2 \le x\le 1+\delta$,
$$h'(x)\ge \frac{2(x-1)\kappa^2 - d^2}{x^3} \cdot p^{-1} \tr\left[\left([(X^{(2)})^\top X^{(2)}]^{-1} + x^{-2} [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right] \gtrsim p\left(d^2 + |x-1|\kappa^2\right).$$
Thus we get that if $1+2d^2/\kappa^2 +p^{-1/4+c}\le x\le 1+\delta$,
\begin{align*}
h(x)-h(1)\ge h(x)-h(1+d^2/\kappa^2) \ge \int_{1+d^2/\kappa^2}^x h'(x)\dd x \gtrsim p\kappa^2 \cdot |x-1|^2  \gtrsim p\frac{d^4}{\kappa^2} + p^{-1/2+2c}\cdot p\kappa^2 .
\end{align*}
On the other hand, by Claim \ref{claim_largedev1} we have that
$$ g(x)=h(x)+\sigma^2 (n_1+n_2-p)+\OO \left(p^{1/2+c}\left(\sigma^2 +\kappa^2+d^2 \right)\right)$$
with high probability. Combining the above two estimates, we obtain that under the conditions $\sigma^2 \lesssim  \kappa^2$ and $d^2 \le p^{-c_{\varphi}}{\kappa^2}$,
$$g(x)-g(1) = h(x)-h(1) + \OO \left(p^{1/2+c}\left(\sigma^2 +\kappa^2+d^2 \right)\right) >0$$
with high probability, i.e. $x$ cannot be the minimizer of $g$. 
%Then if we choose the constant $c$ in Claim \ref{claim_largedev1} such that $e_1<2e_2$, we have 
%$$ h(x)-h(1)\gtrsim  p^{-1/2+2e_2}\cdot p\kappa^2 \gg p^{-1/2+e_1}\cdot p\left(\sigma^2 +\kappa^2+d^2 \right),$$ where we use the assumption that $\sigma^2$ and $d^2$ are both at most $\OO(\kappa^2)$, which implies the inequality \eqref{gA>g1}. 
Second, using equations \eqref{op_uplow} and  \eqref{h'(A)2}, we obtain that for $1-\delta \le x \le 1-d^2/\kappa^2$,
$$- h'(x)\ge  [2(1-x)\kappa^2 - xd^2] \cdot p^{-1} \tr\left[\left(x^2[(X^{(2)})^\top X^{(2)}]^{-1} + [(X^{(1)})^\top X^{(1)}]^{-1}\right)^{-1}\right]  \gtrsim p\left(d^2 + |x-1|\kappa^2\right).$$
Then with a similar argument as above, we can get that if $1-2d^2/\kappa^2 - p^{-1/4+c}\le x\le 1-\delta$, then 
$$ h(x)-h(1) \gtrsim p\frac{d^4}{\kappa^2} +p^{1/2+2c} \kappa^2, $$
%under equation \eqref{choiceofpara0} \HZ{I don't see which condition we want to use here?}, 
which again implies that $g(x)>g(1)$, i.e. $x$ cannot be the minimizer of $g$. In sum, we obtain that the minimizer $\hat x$ must satisfy equation \eqref{hatw_add1}.
%Next we prove the following estimate on the optimizer $\hat x$: with high probability,
%%\begin{lemma}
%%For the isotropic model, we have
%\be\label{hatv_add1}
%|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2 + \e_0 /2+ 2\e}.
%\ee
%%\end{lemma}
%%\begin{proof}
% In fact, from the proof of Claim \ref{lem_hat_v} above, one can check that if $C\cal E \le |w-1| \le 2C\cal E$ for a large enough constant $C>1$, then $|h'(w)|\gtrsim \sigma^2$. Moreover, under equation \eqref{choiceofpara} we have
%$$h(w) =\OO(\sigma^2),\quad \text{for}\quad   |w-1|\le 2C\cal E.$$
%Thus we obtain that for $|w-1|\ge 2C\cal E$,
%$$\left|h(w) - h(\hat w)\right|\ge |h(w)-\min\{h(w_1),h(w_2)\}|\gtrsim \sigma^2 \cal E \gtrsim \cal E \cdot h(\hat w),$$
%which leads to $g(w) > g(\hat w)$ with high probability by equation \eqref{boundv-w}. Thus $w$ cannot be a minimizer of $g(v)$, and we must have $|\hat v - 1|\le 2C\cal E$. %Together with equation \eqref{hatw_add1}, we conclude equation \eqref{hatv_add1}.
%%\end{proof}
%
%{\cor require $\kappa^2\gg d^2 + p^{-1/2+c}\sigma^2$}
\end{proof}

\paragraph{Part $3$: Applying Theorem \ref{thm_main_RMT}.}
We provide the predication loss using the optimal $\hat B$ in equation \eqref{eq_Bhat_2task} and optimal $\hat x=\hat x_1/\hat x_2$:
\begin{align}
L(\hat{\beta}_2^{\MTL}) &=\left\|\Sigma_2^{1/2} \left( \hat B \hat x_2 - \beta_2\right)\right\| \nonumber\\
&=  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat x)^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat x\beta_1-\hat x^2\beta_2)+ (X^{(2)})^\top \epsilon^{(2)} + \hat x   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2.\label{Lbeta_HPS}
\end{align}
%where we denoted $$\hat \Sigma:= \hat x^2 (X^{(1)})^\top X^{(1)}  + (X^{(2)})^\top X^{(2)} .$$
To illustrate the idea of bias-variance decomposition, we assume that $\hat x$ is independent of $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$ for now. Then we can calculate the partial expectation of $L(\hat{\beta}_2^{\MTL})$ over the randomness in $\epsilon_1$ and $\epsilon_2$:
 \begin{align}
	\exarg{\epsilon^{(1)}, \epsilon^{(2)}}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2,\beta_1,\beta_2}  &=  \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} (\hat x \beta_1 - \hat x^2\beta_2)}^2 \nonumber\\ %\label{eq_bias_2task} 
			& + \sigma^2\cdot \bigtr{ X^{(2)} \hat\Sigma(\hat x)^{-1}\Sigma^{(2)}\hat\Sigma(\hat x)^{-1} (X^{(2)})^\top} \nonumber\\
			&+ \sigma^2\cdot \hat x^2 \bigtr{ X^{(1)} \hat\Sigma(\hat x)^{-1}\Sigma^{(2)}\hat\Sigma(\hat x)^{-1} (X^{(1)})^\top}\nonumber\\
			%+\sigma^2\cdot \bigtr{\Sigma_2(\frac{W_1^2}{W_2^2} X_1^{\top}X_1 + X_2^{\top}X_2)^{-1} } \nonumber\\ 
			&= \bignorm{(\Sigma^{(2)})^{1/2}\hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} (\hat x \beta_1 - \hat x^2\beta_2)}^2 +\sigma^2  \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).\nonumber
			%\label{eq_var_2task}
\end{align}
This gives a bias-variance decomposition of the MTL prediction loss: the first term on the right-hand side is the bias term, while the second term is the variance term. For the random-effects model, we further take the partial expectation over $\beta_1$ and $\beta_2$: 
 \begin{align}
	& \exarg{\epsilon^{(1)}, \epsilon^{(2)},\beta_1,\beta_2}{L(\hat{\beta}_2^{\MTL}) \mid X_1, X_2} \nonumber\\
	&	=  \hat x^2 \left[(\hat x-1)^2 \kappa^2 + \frac{(\hat x^2+1)d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma(\hat x)^{-1} \Sigma^{(2)} \hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} }\nonumber \\
	&+\sigma^2 \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).
			\label{eq_var_2task}
\end{align}
One can compare the above expression with the predication loss for STL, that is,
\begin{align}
	\exarg{\epsilon_2}{L(\hat{\beta}_2^{\STL}) \mid X_2} = \sigma^2 \cdot \bigtr{\Sigma^{(2)} (X_2^{\top} X_2)^{-1}}, \label{eq_var_stl}
\end{align}
we observe that while the bias of $\hat{\beta}_2^{\MTL}$ is always larger than that of $\hat{\beta}_2^{\STL}$, which is zero, the variance of $\hat{\beta}_2^{\MTL}$ is always lower than that of $\hat{\beta}_2^{\STL}$.\footnote{To see why this is true, we apply the Woodbury matrix identity over the second term on the right-hand side of equation \eqref{eq_var_2task} and use the fact that for the product of two PSD matrices, its trace is always nonnegative.}
In other words, training both tasks together helps predict the target task by reducing variance while incurring a bias.
Therefore, whether multi-task learning outperforms single-task learning is determined by the bias-variance decomposition!

@@

In general, for $\hat x$ as  the minimizer of $g(x)$, it depends on $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$. However, using the concentration estimates in Lemma \ref{largedeviation}, we can still obtain the expression in equation \eqref{eq_var_2task} up to a small error. In particular, our proof will not use Lemma \ref{lem_hat_v}.

%control the difference between $g(x)$ and its partial expectation. 
 \begin{claim}\label{claim_largedev2}
 In the setting of Proposition \ref{cor_MTL_loss}, if $\hat x=\OO(1)$, then we have that with high probability,
$$\left| L(\hat{\beta}_2^{\MTL})- \cal L(\hat x)\right| \le p^{-1/2+c} \left(\sigma^2 +\kappa^2 +d^2 \right),$$
for any small constant $c>0$. Here we introduced the function
	\cal L(\hat x)	:=  &\hat x^2\left[(\hat x-1)^2 \kappa^2 + \frac{(\hat x^2+1)d^2}2\right] \cdot p^{-1}\bigtr{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma(\hat x)^{-1} \Sigma^{(2)} \hat\Sigma(\hat x)^{-1} (X^{(1)})^{\top}X^{(1)} } \\
	& +\sigma^2\cdot \tr\left(\Sigma^{(2)}\hat \Sigma(\hat x)^{-1}\right).
 The proof of this claim is very similar to the one for Claim \ref{claim_largedev1}, where the only difference is that $\hat x$ now may depend on $\epsilon^{(1)}$, $\epsilon^{(2)}$, $\beta_1$ and $\beta_2$. In this case, we can still use  Lemma \ref{largedeviation} to conclude the proof because $\hat x$ can be pulled out as a coefficient. Here we only give an outline of the proof without writing down all the details. Recall that $\beta_0$ is the shared component of $\beta_1$ and $\beta_2$ with i.i.d. Gaussian entries of mean zero and variance $p^{-1}\kappa^2$. Moreover, we denote the task-specific components by $\wt\beta_1$ and $\wt\beta_2$, whose entries are i.i.d. Gaussian random variables of mean zero and variance $p^{-1} d^2 / 2$. Then we write $L(\hat{\beta}_2^{\MTL}) $ in \eqref{Lbeta_HPS} as:
L(\hat{\beta}_2^{\MTL})  =&  \left\| (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat x)^{-1} \left[(X^{(1)})^\top X^{(1)} (\hat x -\hat x^2)\beta_0+(X^{(1)})^\top X^{(1)} \hat x\wt \beta_1 - (X^{(1)})^\top X^{(1)}  \hat x^2\wt \beta_2 \right] \right. \nonumber\\
&\left. + (\Sigma^{(2)})^{1/2}\hat \Sigma(\hat x)^{-1}\left[ (X^{(2)})^\top \epsilon^{(2)} + \hat x   (X^{(1)})^\top \epsilon^{(1)} \right]\right\|^2. \label{expand_15}
As in \eqref{expand_6}, we can expand this expression into the sum of 15 terms, and bound each term as in \eqref{eq_hA1}-\eqref{eq_hA56}. For example, for the main term $\hat x^2 \wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(1)})^\top X^{(1)}\wt\beta_1$, using 
Lemma \ref{largedeviation} and \eqref{op_X12}-\eqref{op_Sig1}, we can obtain the following estimate with high probability for any small constant $c>0$:
&  \left|\wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(1)})^\top X^{(1)}\wt\beta_1 - \frac{d^2}{2} \cdot p^{-1}\bigtr{(X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(1)})^\top X^{(1)}}\right| \\
&\le p^{-1+c}d^2 \cdot \left\| (X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(1)})^\top X^{(1)} \right\|_F \\
&\le p^{-1/2+c}d^2 \cdot  \left\| (X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(1)})^\top X^{(1)} \right\|  \lesssim p^{-1/2+c}d^2.
For the cross term $\hat x \wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(2)})^\top  \epsilon^{(2)}$, using Lemma \ref{largedeviation} and \eqref{op_X12}-\eqref{op_Sig1}, we can obtain the following estimate with high probability for any small constant $c>0$:
 \left|\wt \beta_1^\top (X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(2)})^\top  \epsilon^{(2)}\right|  & \le p^c \cdot \sigma\sqrt{p^{-1}d^2} \cdot \|(X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(2)})^\top  \|_F \\
&\lesssim p^{c}\sigma d \cdot  \|(X^{(1)})^\top X^{(1)} \hat \Sigma(\hat x)^{-1}  \Sigma^{(2)} \hat \Sigma(\hat x)^{-1}  (X^{(2)})^\top  \| \\
&\lesssim p^{-1/2+c}\sigma d\le p^{-1/2+c}(\sigma^2+ d^2).
The rest of the terms in the expansion of \eqref{expand_15} can be bounded in the same way, and we omit the details.

In the setting of Proposition \ref{cor_MTL_loss}, we can further simplify $\cal L(\hat x)$ using Claim \ref{lem_hat_v} and $\Sigma^{(1)}=\Sigma^{(2)}=\id_p$.

@@

&\left|\cal L(\hat x)- d^2\cdot {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] -\sigma^2  \tr\left({\hat \Sigma^{-1}  }\right)\right| \nonumber\\
&\lesssim  \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) , \label{simple1}

@@

for any small constant $c>0$, where we recall that $\hat \Sigma=\hat \Sigma(1)$ as defined in \eqref{def hatsig}.

@@

Using Lemma \ref{lem_hat_v} and equations \eqref{op_X12}-\eqref{op_Sig1}, we can obtain that with high probability,

@@

 \|\hat\Sigma^{-1}-\hat\Sigma(\hat x)^{-1}\| &\le |\hat x^2-1|\|\hat\Sigma^{-1}\| \| (X^{(1)})^\top X^{(1)}\|\|\hat\Sigma(\hat x)^{-1}\|  \lesssim p^{-1}\left(\frac{d^2}{\kappa^2} + p^{-1/4+c}\right),\label{est111}

@@

for any small constant $c>0$. With similar arguments, we can get that with high probability,

@@

&\left\|\left(\hat\Sigma^{-2}-\hat\Sigma(\hat x)^{-2}\right)\left((X^{(1)})^\top X^{(1)} \right)^2\right\| \lesssim  \frac{d^2}{\kappa^2} + p^{-1/4+c} .

@@

Moreover, using equations \eqref{op_X12}-\eqref{op_Sig1} we can bound that with high probability,
%$$\bignorm{ (X^{(1)})^{\top}X^{(1)} \hat\Sigma^{-1} \Sigma^{(2)} \hat\Sigma^{-1} (X^{(1)})^{\top}X^{(1)}}\lesssim 1.$$
%which further implies that with high probability,
